\subsection{Learning from demonstration}
\label{sec:lfd}
This section is dedicated to presenting and analyzing various approaches for solving the LfD problem described in Section~\ref{sec:problem_formulation}. Specifically, this section follows the taxonomy derived from studying different review papers~\cite{kaelbling1996reinforcement_survey,argall2009robot_learning_from_demonstration,hussein2017imitation_learning_survey,fang2019survey,zheng2021imitation_progress_taxonomies_opportunities,zare2024survey}. Figure~\ref{fig:il_taxonomy} provides a graphical representation of the proposed taxonomy. The methods are first categorized based on the type of demonstration, either \textit{State-Action} or \textit{State-only}, followed by an overview of the different methodologies. The proposed taxonomy highlights the learning algorithm and main components for each methodology.

This chapter is divided into the following sections. Section~\ref{sec:bc} reviews methods for the fully-supervised learning methodology known as Behavioral Cloning. 
\newline Section~\ref{sec:irl} discusses methods under the umbrella of Inverse Reinforcement Learning, which solve the inverse optimization problem by first learning the reward function and then using it to guide policy optimization following a reinforcement paradigm
\newline Section~\ref{sec:gail} reviews methods leveraging the concept of \textit{Generative Adversarial Learning} to optimize policy parameters and learn demonstrated behaviors, classified as Generative Adversarial Imitation Learning methods.
\newline Finally, Section~\ref{sec:lfo} covers the most recent methodology, Learning from Observation, characterized by optimizing the learner policy using state-only demonstrations.
\input{figures/ch1/taxonomy.tex}
Each paragraph will present the research in the same way, i.e., the proposed works will be presented in a temporal order, highlighting the evolution of techniques and approaches over time.
It is important to note that with respect to the problem managed in this thesis, the most relevant and most related literature is the one preseneted in Section \ref{sec:bc}. However, for sake of clarity and completeness, all the other approaches will be described with their pros/cons, with some considerations with respect to the proposal of this thesis. \smalltodo{ADD THIS DISCUSSION}.


\input{chapters/ch1/sota/bc.tex}
\input{chapters/ch1/sota/irl.tex}
\input{chapters/ch1/sota/gail.tex}
\input{chapters/ch1/sota/lfo}




